Modular Synthesis of Verified Verifiers of Computation with STV Algorithms

### 2019-03-14 paper/paper.tex
typos all corrected

We observe that every instance of STV performs the same
\emph{actions} (e.g. count first preferences, transfer a surplus,
eliminate a candidate, etc.) where the precise interpretation of
actions differs. We formalise these actions as rules that advance the state of the count towards a final result. Evidence for an instance of computation then becomes a sequence of rule applications applied to perform this computation. While generic STV abstracts the commonalities, each \emph{instance} of STV formally specifies the
pre- and postconditions for each rule as well as a boolean function that decides whether a rule has been applied correctly.

Executable evidence verifiers are constructed almost
automatically:  nearly all of the ensuing proof obligations are discharged generically, and we fully  automate the translation into machine-level verified executable using CakeML.

@@

In summary, we present a framework that allows us to formally specify instances of STV and sythesise an executable certificate verifier in a modular way that is itself verified down to the machine level.
\section{Evidences}
\label{sec:DataEv}

@@

The evidence as such must be enough informative to provide transparency of tallying and also allow  voters, or at least a large pool of scrutineers, to verify for themselves that the tally is authentically processed as per instructions of the counting scheme  so that winners truly reflect will of the voters. Therefore the question becomes what information is enough  for establishing transparency and verifiability of tallying  and what data structure for recording the information in the evidence we should choose.
\subsection{Characterisation of Evidence} 
A comprehensive analysis of different STV algorithms reveals existence of a common underlying abstract data type among STV algorithms as follows. 
%Each STV scheme operates on the following data.
\begin{itemize}
\item Some ballots to deal with (assuming) recorded as cast by voters. Each ballot is a pair consisting of a list expressing the preference of a voter and a fractional transfer value,
\item a quota which is the least amount of votes needed to be elected, 
\item the candidates competing in the election, 
\item the number of vacancies for which candidates compete,
\item the \emph{discrete states of computation} which encapsulate necessary  pieces of information needed to transparently know how the tallying has progressed from beginning to end and to be able to independently verify its correctness provided access to the information is granted. 
\end{itemize}
There are two kinds of discrete states of computation. Some of them are final states where tallying has reached a halting state by finding the winners of the election. The other consist of non-final ones where bits and pieces of data are manipulated according to counting mechanism of the STV scheme used, to eventually obtain an end result. Each non-final state comprises seven pieces of data: 
\begin{itemize}
 \item\textbf{uncounted ballots} which await being counted,
 \item\textbf{tally} for holding information about the amount of votes that each candidate has hitherto attracted %up to this state of computation,
 \item\textbf{pile} to know the ballots allocated to each candidate, 
 \item\textbf{backlog of elected} candidates who have exceeded the quota and their surplus votes awaits being transferred,
 \item\textbf{backlog of eliminated} ones whose votes awaits transfer,
 \item\textbf{elected candidates} who are the already elected ones,% whose surplus votes must be transferred,
 \item\textbf{continuing candidates} still continuing in the election. 
 \end{itemize}
The non-final and final states specified above provide us with all information one needs. We therefore find a perfect characterisation for evidence. 
\begin{definition}\label{formalEv}
Suppose an STV algorithm $\mathcal{A}$ is given and $Q$, $s$ and $C$
are respectively the quota, number of vacancies and competing
candidates in an election whose counting algorithm is $\mathcal{A}$. Also assume  $y$ is the output of an execution of $\mathcal{A}$ on an input $x$. By evidence $\hat{\omega}$ for the execution  we mean the quadruple $(Q,s,C,\Omega)$, where $\Omega$ is the chronological sequence of all states of this instance of computation visited from the initial state where $x$ is input to the final state where $y$ is output.   
\end{definition}
\section{The Generic STV  Machine}
\label{sec:Machine}
\setlength{\abovedisplayskip}{0.5em}
\setlength{\belowdisplayskip}{0.5em}
We have already discussed an abstract data type whose values
formally represent evidence. However we do not yet possess an
operational semantics for manipulating data to validate evidence.
On one hand, the semantics must be flexible enough to accommodate
variations existing in the counting mechanism of STV algorithms so
that we can produce verifiers  specifically operating for validation
of evidence output by their associated scheme. On the other hand it
must facilitate designing and developing the framework modularly
while offering automation for synthesising verifiers.

@@

The underlying abstract data type and the universal algorithmic
pattern found in STV characterise components of a finite state
machine which we call \emph{generic
STV}~(Definition\ref{STVMachine}). The set of machine states consist
of the discrete states of computation $\mathcal{S}$ mentioned
earlier. The seven generic actions given in
Definition~\ref{transitions} form the transition labels
$\mathcal{T}$ of the machine. Last, the everywhere present pre- and
post-conditions which define (content of)  the actions, and
constitute a small-step operational semantics for the machine.  
Note that the states of computation recorded as evidence now become identical with machine states.

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ 
 where $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then the \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$.  
\end{definition}
There are nonetheless variations in STV algorithms  which we  recognise  by separate \emph{instantiations} into the machine.  
\subsection{Instantiations of the Machine with Specific STV schemes}
%\noindent
Differences exist between counting algorithms of STV schemes.  For
example in the STV used for electing the upper house representatives
of the Australian State of Victoria (Victoria STV)~\cite{vic},
distributing votes of an eliminated candidate occurs step by step in
order of the magnitude of the fractional values that those votes
carry.  Consequently, a pre-condition for applying the action is to
first assure that upon each instance of applying transfer-removed
action (a) exactly those votes which have the same fractional value
are transferred and (b) the transfer happens according to the
magnitude order. In contrast, for example, the STV used in the lower
house elections of the Australian state of Tasmania (Tasmania STV)
does not have such precondition simply because votes of an
eliminated candidate are transferred in a single application of
transfer-removed. There are also variations in postconditions. For
example transfer-elected of the ACT STV~\cite{act} used in the lower
house elections of the Australian Capital Territory  requires distributing only the last parcel of surplus votes received by an elected candidate. But Victoria STV specifies transferring all surplus votes of an elected candidate.

@@

The operational semantics of each instantiation allows  to perform operations such as validation on evidence. Definition~\ref{verifier} lays down what verifying an evidence amounts to. Informally speaking, to check if  evidence is valid one simply needs to inspect if transitions  between each two consecutive states of computation appearing in the evidence occur by a legitimate application of a counting action of the scheme used.

@@

First preferences for each candidate are computed. Ballots counted in favour of particular candidates are placed onto that candidate’s pile. In this election, $a$ is the first preference on $b_{1}$, $b_{2}$, and $b_{3}$, $b$ receives $b_{4}$, and $c$ receives $b_{5}$. Tallies are updated so that tally of $a$ becomes 3, and $b$
and $c$ each reach $1/1$. As $a$ exceeds the quota, (s)he is elected and the fractional value of his surplus is updated according to the ACT scheme. Then $a$'s surplus is transferred to continuing candidates $b$ and $c$ and votes are counted according to next preferences shown on $a$'s surplus votes. Candidate $b$ and $c$ each attract one vote. However  neither of them reaches to the quota. Hence the weakest one $b$ is eliminated. As the number of elected and continuing candidates is equal to 
%does not exceed 
the initial vacancies, the election terminates by declaring $a$ and
$c$ to be winners.
\begin{small} 
\begin{figure}[b]
\begin{tabular}{c@{\hspace{2cm}}c} 
\AxiomC{\tiny 8/3} \noLine

@@

\section{The Framework Architecture}\label{sec:Arc} 
%\noindent
We progress through three macro-level phases to develop the framework. For the first phase we use HOL4 to formally specify, implement and verify the generic STV, its instantiations, verifiers, and every auxiliary assertions needed for these ends. The second step relies on the verified proof translation tool of CakeML to provably correctly translate implementations of the first phase into equivalent CakeML implementations. The last phase uses the ecosystem of CakeML and its verified compiler to generate machine executable evidence verifiers.

@@

\paragraph*{\textbf{Auxiliary}} The module contains specification, implementation and verification of the generic STV. It consists of subdivisions for formal specification of the machine and its components, implementation of functions  meant to be the computational counterparts of the specifications and verification of the implementations by proving that they logically satisfy their respective specifications. Moreover we include specification, implementation and verification of helper assertions which appear in many well-known STV schemes here so that Auxiliary also serves as a comprehensive library for STV. 
\paragraph*{\textbf{Instantiations}}
  Instantiations of the generic STV happen in separate modules which we have schematically shown only two, namely   \textbf{STV}$_{1}$ and \textbf{STV}$_{2}$. An instantiation module \textbf{STV}$_{i}$ for an algorithm $\mathcal{A}$ consists of a subdivision for specification $\hat{\mathcal{A}}_{spec}$ of  the scheme and another one its implementation $\hat{\mathcal{A}}_{dec}$. 
\paragraph*{\textbf{Proofs}} We prove that for an algorithm $\mathcal{A}$ in the above modules,  $\hat{\mathcal{A}}_{dec}$ is logically equivalent to $\hat{\mathcal{A}}_{spec}$. We automate proofs so that they work for various instantiations. 
\paragraph*{\textbf{Verifier}} This module formally realises  the notion of verifier in Definition~\ref{verifier}. However, it is developed in a way that automatically operates  regardless of which STV algorithm is instantiated into the machine. It has three subdivisions. A  specification $\hat{\mathcal{V}}_{spec}$ defining what a verifier is, the implementation $\hat{\mathcal{V}}_{dec}$ and proofs of equivalence of $\hat{\mathcal{V}}_{spec}$ and $\hat{\mathcal{V}}_{dec}$.
\paragraph*{\textbf{Translation}} We translate each implementation $\hat{f}_{imp}$ in HOL4 to a proven  equivalent implementation $f_{\tau}$ in CakeML's environment by using the verified proof translator of CakeML. For example, an implementation  $\hat{\mathcal{A}}_{imp}$ of a scheme $\mathcal{A}$ and its   verifier $\hat{\mathcal{V}}_{imp}$ are  respectively translated into $\mathcal{A}_{\tau}$  and $\mathcal{V}_{\tau}$. 
\paragraph*{\textbf{DeepSpec}} 
For any translated verifier $\mathcal{V}_{\tau}$ and the translated evidence parser $\mathcal{P}_{\tau}$, we obtain logically equivalent deep CakeML embeddings  $\mathcal{V}^{*}$ and $\mathcal{P}^{*}$, respectively. We then use them to establish end-to-end desired properties, based on CakeML's I/O model, about the interaction of the executable verifier with its hosting operating system.   
\tikzset{
    state/.style={
           rectangle,
           rounded corners,
           draw=black, 
           minimum height=2em,
           inner sep=2pt,
           text centered,
           },
}
\begin{figure}[t]
\centering
%\framebox[\columnwidth]{
\begin{tikzpicture}[->,>=stealth']

@@

\end{tikzpicture}
\caption{Architecture of the Framework, direction of arrow represents module dependencies}
\label{PicArc}
\end{figure}
\paragraph*{\textbf{Compilation}} 
Instantiations of CakeML's  compiler for generating machine executable verifiers happen in this module. For any deeply embedded $\mathcal{V}^{*}_{i}$, corresponding to $\hat{\mathcal{V}}_{dec}$ in \textbf{STV}$_{i}$, using proofs  established in \textbf{DeepSpec}, we instantiate the compiler to synthesise executable  verifier $\mathcal{V}_{i}$.  
\section{Generic STV Formalised}\label{sec:GenCertVer}
We proceed through three steps, namely specification, implementation and verification of implementations in HOL4 to formally verify the machine and its components.
\paragraph*{Specification}  
Recall that we require minimising the trusted computing base  for computing with generated verifiers. The specification module consists of properties which we can use to verify that our implementations of  STV algorithms in HOL4 indeed match with the description of the counting methods used.  
 In light of Definition~\ref{STVMachine} these properties   actually form the semantics of the generic STV machine.   
\paragraph*{Implementation} Our final objective for creating the
framework is to produce means for validating concrete items of evidence. To this end, we need decision procedures used for deciding whether or not a given evidence is valid. Therefore we define boolean-valued functions in the implementation sub-module to computationally realise logical declarations in the specification module. In particular, we implement decision procedures whose computational content  provably realises the specification of the machine semantics.  
\paragraph*{Verification} Here we formally prove that the implementations meet the expectations of their respective specifications. Therefore through verification we come to eliminate a trusted layer  required to lay in our framework.

@@

\item As the abstract syntactic representation of evidence closely represents concrete evidences, designing the parser and inspecting  its  correctness is less challenging.  
 \item HOL4 has well developed libraries containing verified
 assertions of operations on lists. By using lists to represent the

@@

 tactics for discharging proof obligations on assertions involving
 list structure and operations on it. This significantly eases the
 process of verification. 

@@

There is also a reason for choosing the type $rat$  
%as the underlying arithmetic for the formalisation, 
instead of e.g. floating numbers. Elections with an STV counting scheme have a small margin of victory especially for the last vacancy left to fill \cite{MBlo}. This margin may happen to be less than magnitude of the error caused by accumulation of rounding errors as it is with floating points.  Hence one must sensibly avoid the high  cost of electing a wrong candidate by falling into traps of imprecise calculations due to unintelligent choices. Using exact fractions allows safe handling of calculations. 
\subsection{The Semantics and its Auxiliary Components}
\label{sec:MachineSem}

The semantics of each machine transition label consists of conjunctions of formally specified general pre- and post-conditions across STV schemes which enforce when and how to take a tallying action and what the immediate effect is. To demonstrate the process of specification, implementation and verification of the machine transitions and their semantics, we discuss the transfer-elected transition.

@@

%\footnote{The Tasmania STV also uses similar surplus transfer mechanism.}.
 Instantiation of other machine transitions proceed in a  similar manner. Under section 'Step 3' and 'transfer surplus from elected candidates' the protocol explains under what conditions and how to distribute surplus votes. We summarise and rephrase these sections as follows. 

@@

\item[$\bullet_{6}$] only the last parcel of votes  received (which resulted in a surplus) is transferred. It may be that the last parcel is the only parcel in a candidate's pile (if only one application of the count action has occurred), or more parcels exist (if several actions of count has happened as a result of earlier  elect, transfer or eliminate actions)

@@

The lines denoted by the star symbol~($\star$) in TransferActSpec formalise the Clause~$\bullet_{7}$ above. Also note that we place the last parcel of the pile of the candidate $c$ whose votes are transferred first into the list of uncounted ballots so that in the subsequent transition count deals with distributing the votes and ``recalculating'' candidates' tallies and piles. Finally, because of efficiency purposes our system is designed to update the fractional transfer value of the surplus in formalisation of the elect transition instead of the transfer-elected transition.

@@

For example, we represented instantiation of transfer-elected with

@@

work. This is different in the actual proof engineering development.  We carry out different instantiations of the machine in separate modules but uniformly name instantiations of the transitions. For example, for instantiation of the machine with the ACT STV, we have a module accordingly named and inside the module we refer, for instance, to instantiation of transfer-elected as TransferDec or TransferSpec without the infix Act.

@@

\emph{Scenario three.} 
 Suppose there is a clause in description of $\mathcal{A}$ whose specification, and  therefore implementation,  does not exist in the auxiliary module. Then one trivially has to extend the auxiliary module by specifying and implementing that clause and then verifying the implementation correct against the specification.  If one was overconfident in their implementation, then they can take the implementation as its specification in which case  no verification is required.  Note that as including the generic formal machine transitions in the semantics of each transition instantiation is mandatory  
 there are few  numbers of such clauses and consequently few lines of encoding needed to formalise them.

@@

 The function check\textunderscore{}count accepts a file as input on the command line, opens the file consisting of  evidence which we intend to validate, parses the header of the evidence consisting of the quota, number of seats and competing candidates in the election. If the header is well-formed, then check\textunderscore{}count proceeds to parse two judgement lines at a time and if the lines  successfully parse into  values of the type judgement (the type of the machine states) then checks if  the transition happening between them \emph{corresponds with the specification of the verifier in HOL4 ($\mathcal{V}_{spec}$)}. This process of reading evidence lines, parsing and checking them continues until either the evidence is accepted as valid or  check\textunderscore{}count encounters a malformed judgement line or an invalid transition step from one parsed judgement (machine state) to another  in which case it returns an error messages  informing us where it occurs.

@@

Use of theorem proving to verify vote counting in elections
is an emerging field. The only use of theorem proving and certifying algorithms~\cite{CertAlg} that we are aware is our own previous work~\cite{GhaleVSTTE} which just verifies a single instance of STV. While the overall approach is similar, the underlying data structure and the algorithmic content which is being formalised does not allow modular specification, verification and synthesis of verifiers for different types of STV: only a single STV algorithm is
being verified.  Adapting to other schemes
requires considerable effort of re-encoding (sometimes the whole data structure) and non-trivial expertise with HOL4 and CakeML.

Other approaches focus on the  formalisation and verification of algorithms for correctly computing election results, rather than checking correctness of computation performed. This is usually based on so-called heavy weight formal methods (interactive theorem proving) and uses a variety of 
%different
 tools.

>>>>

We observe that every instance of STV performs the same
\emph{actions} (e.g. count first preferences, transfer a surplus,
eliminate a candidate, etc.) where the precise interpretation of
actions differ. We formalise these actions as rules that advance the state of the count towards a final result. Evidence for an instance of computation then becomes a sequence of rule applications applied to perform this computation. While generic STV abstracts the commonalities, each \emph{instance} of STV formally specifies the
pre- and postconditions for each rule as well as a boolean function that decides whether a rule has been applied correctly.

Executable evidence verifiers are constructed almost
automatically:  nearly all of the ensuing proof obligations are discharged generically, and we fully automate the translation into a machine-level verified executable using CakeML.

@@

In summary, we present a framework that allows us to formally specify instances of STV and synthesise an executable certificate verifier in a modular way that is itself verified down to the machine level.
\section{Evidences}
\label{sec:DataEv}

@@

The evidence as such must be enough informative to provide transparency of tallying and also allow voters, or at least a large pool of scrutineers, to verify for themselves that the tally is authentically processed as per instructions of the counting scheme so that winners truly reflect the will of the voters. Therefore the question becomes what information is enough for establishing transparency and verifiability of tallying and what data structure for recording the information in the evidence we should choose.
\subsection{Characterisation of Evidence} 
A comprehensive analysis of different STV algorithms reveals the existence of a common underlying abstract data type among STV algorithms as follows. 
%Each STV scheme operates on the following data.
\begin{itemize}
\item Some ballots to deal with (assuming) recorded as cast by voters. Each ballot is a pair consisting of a list expressing the preference of a voter and a fractional transfer value,
\item a quota which is the least amount of votes needed to be elected, 
\item the candidates competing in the election, 
\item the number of vacancies for which candidates compete,
\item the \emph{discrete states of computation} which encapsulate necessary pieces of information needed to transparently know how the tallying has progressed from beginning to end and to be able to independently verify its correctness provided access to the information is granted. 
\end{itemize}
There are two kinds of discrete states of computation. Some of them are final states where tallying has reached a halting state by finding the winners of the election. The other consist of non-final ones where bits and pieces of data are manipulated according to the counting mechanism of the STV scheme used, to eventually obtain an end result. Each non-final state comprises seven pieces of data: 
\begin{itemize}
 \item\textbf{uncounted ballots} which await being counted,
 \item\textbf{tally} for holding information about the amount of votes that each candidate has hitherto attracted %up to this state of computation,
 \item\textbf{pile} to know the ballots allocated to each candidate, 
 \item\textbf{backlog of elected} candidates who have exceeded the quota and their surplus votes awaits being transferred,
 \item\textbf{backlog of eliminated} ones whose votes await transfer,
 \item\textbf{elected candidates} who are the already elected ones,% whose surplus votes must be transferred,
 \item\textbf{continuing candidates} still continuing in the election. 
 \end{itemize}
The non-final and final states specified above provide us with all the information one needs. We therefore find a perfect characterisation for evidence. 
\begin{definition}\label{formalEv}
Suppose an STV algorithm $\mathcal{A}$ is given and $Q$, $s$ and $C$
are respectively the quota, number of vacancies and competing
candidates in an election whose counting algorithm is $\mathcal{A}$. Also assume  $y$ is the output of an execution of $\mathcal{A}$ on an input $x$. By evidence $\hat{\omega}$ for the execution we mean the quadruple $(Q,s,C,\Omega)$, where $\Omega$ is the chronological sequence of all states of this instance of computation visited from the initial state where $x$ is input to the final state where $y$ is output.   
\end{definition}
\section{The Generic STV  Machine}
\label{sec:Machine}
\setlength{\abovedisplayskip}{0.5em}
\setlength{\belowdisplayskip}{0.5em}
We have already discussed an abstract data type whose values
formally represent evidence. However we do not yet possess an
operational semantics for manipulating data to validate evidence.
On one hand, the semantics must be flexible enough to accommodate
variations existing in the counting mechanism of STV algorithms so
that we can produce verifiers  specifically operating for validation
of evidence output by their associated scheme. On the other hand it
must facilitate designing and developing the framework modularly
while offering automation for synthesising verifiers.

@@

The underlying abstract data type and the universal algorithmic
pattern found in STV characterise components of a finite state
machine which we call \emph{generic
STV}~(Definition\ref{STVMachine}). The set of machine states consist
of the discrete states of computation $\mathcal{S}$ mentioned
earlier. The seven generic actions in
Definition~\ref{transitions} form the transition labels
$\mathcal{T}$ of the machine. Last, the everywhere present pre- and
post-conditions which define (content of)  the actions, and
constitute a small-step operational semantics for the machine.  
Note that the states of computation recorded as evidence now become identical with machine states.

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ 
 where $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then the \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$.  
\end{definition}
There are nonetheless variations in STV algorithms  which we  recognise  by separate \emph{instantiations} into the machine.  
\subsection{Instantiations of the Machine with Specific STV schemes}
%\noindent
Differences exist between counting algorithms of STV schemes.  For
example in the STV used for electing the upper house representatives
of the Australian State of Victoria (Victoria STV)~\cite{vic},
distributing votes of an eliminated candidate occurs step by step in
order of the magnitude of the fractional values that those votes
carry.  Consequently, a precondition for applying the action is to
first assure that upon each instance of applying transfer-removed
action (a) exactly those votes which have the same fractional value
are transferred and (b) the transfer happens according to the
magnitude order. In contrast, for example, the STV used in the lower
house elections of the Australian state of Tasmania (Tasmania STV)
does not have such precondition simply because votes of an
eliminated candidate are transferred in a single application of
transfer-removed. There are also variations in postconditions. For
example transfer-elected of the ACT STV~\cite{act} used in the lower
house elections of the Australian Capital Territory requires distributing only the last parcel of surplus votes received by an elected candidate. But Victoria STV specifies transferring all surplus votes of an elected candidate.

@@

The operational semantics of each instantiation allows performing operations such as validation on evidence. Definition~\ref{verifier} lays down what verifying evidence amounts to. Informally speaking, to check if evidence is valid one simply needs to inspect if transitions between each two consecutive states of computation appearing in the evidence occur by a legitimate application of a counting action of the scheme used.

@@

First preferences for each candidate are computed. Ballots counted in favour of particular candidates are placed onto that candidate’s pile. In this election, $a$ is the first preference on $b_{1}$, $b_{2}$, and $b_{3}$, $b$ receives $b_{4}$, and $c$ receives $b_{5}$. Tallies are updated so that tally of $a$ becomes 3, and $b$
and $c$ each reach $1/1$. As $a$ exceeds the quota, (s)he is elected and the fractional value of his surplus is updated according to the ACT scheme. Then $a$'s surplus is transferred to continuing candidates $b$ and $c$ and votes are counted according to next preferences shown on $a$'s surplus votes. Candidate $b$ and $c$ each attract one vote. However neither of them reaches the quota. Hence the weakest one $b$ is eliminated. As the number of elected and continuing candidates is equal to 
%does not exceed 
the initial vacancies, the election terminates by declaring $a$ and
$c$ to be winners.
\begin{small} 
\begin{figure}[b]
\begin{tabular}{c@{\hspace{2cm}}c} 
\AxiomC{\tiny 8/3} \noLine

@@

\section{The Framework Architecture}\label{sec:Arc} 
%\noindent
We progress through three macro-level phases to develop the framework. For the first phase, we use HOL4 to formally specify, implement and verify the generic STV, its instantiations, verifiers, and every auxiliary assertion needed for these ends. The second step relies on the verified proof translation tool of CakeML to provably correctly translate implementations of the first phase into equivalent CakeML implementations. The last phase uses the ecosystem of CakeML and its verified compiler to generate machine executable evidence verifiers.

@@

\paragraph*{\textbf{Auxiliary}} The module contains specification, implementation and verification of the generic STV. It consists of subdivisions for the formal specification of the machine and its components, implementation of functions meant to be the computational counterparts of the specifications and verification of the implementations by proving that they logically satisfy their respective specifications. Moreover we include specification, implementation and verification of helper assertions which appear in many well-known STV schemes here so that Auxiliary also serves as a comprehensive library for STV. 
\paragraph*{\textbf{Instantiations}}
  Instantiations of the generic STV happen in separate modules which we have schematically shown only two, namely   \textbf{STV}$_{1}$ and \textbf{STV}$_{2}$. An instantiation module \textbf{STV}$_{i}$ for an algorithm $\mathcal{A}$ consists of a subdivision for specification $\hat{\mathcal{A}}_{spec}$ of  the scheme and another one its implementation $\hat{\mathcal{A}}_{dec}$. 
\paragraph*{\textbf{Proofs}} We prove that for an algorithm $\mathcal{A}$ in the above modules,  $\hat{\mathcal{A}}_{dec}$ is logically equivalent to $\hat{\mathcal{A}}_{spec}$. We automate proofs so that they work for various instantiations. 
\paragraph*{\textbf{Verifier}} This module formally realises  the notion of verifier in Definition~\ref{verifier}. However, it is developed in a way that automatically operates  regardless of which STV algorithm is instantiated into the machine. It has three subdivisions. A  specification $\hat{\mathcal{V}}_{spec}$ defining what a verifier is, the implementation $\hat{\mathcal{V}}_{dec}$ and proofs of equivalence of $\hat{\mathcal{V}}_{spec}$ and $\hat{\mathcal{V}}_{dec}$.
\paragraph*{\textbf{Translation}} We translate each implementation $\hat{f}_{imp}$ in HOL4 to a proven  equivalent implementation $f_{\tau}$ in CakeML's environment by using the verified proof translator of CakeML. For example, an implementation  $\hat{\mathcal{A}}_{imp}$ of a scheme $\mathcal{A}$ and its   verifier $\hat{\mathcal{V}}_{imp}$ are  respectively translated into $\mathcal{A}_{\tau}$  and $\mathcal{V}_{\tau}$. 
\paragraph*{\textbf{DeepSpec}} 
For any translated verifier $\mathcal{V}_{\tau}$ and the translated evidence parser $\mathcal{P}_{\tau}$, we obtain logically equivalent deep CakeML embeddings  $\mathcal{V}^{*}$ and $\mathcal{P}^{*}$, respectively. We then use them to establish end-to-end desired properties, based on CakeML's I/O model, about the interaction of the executable verifier with its hosting operating system.   
\tikzset{
    state/.style={
           rectangle,
           rounded corners,
           draw=black, 
           minimum height=2em,
           inner sep=2pt,
           text centered,
           },
}
\begin{figure}[t]
\centering
%\framebox[\columnwidth]{
\begin{tikzpicture}[->,>=stealth']

@@

\end{tikzpicture}
\caption{Architecture of the Framework, direction of arrows represent module dependencies}
\label{PicArc}
\end{figure}
\paragraph*{\textbf{Compilation}} 
Instantiations of CakeML's  compiler for generating machine executable verifiers happen in this module. For any deeply embedded $\mathcal{V}^{*}_{i}$, corresponding to $\hat{\mathcal{V}}_{dec}$ in \textbf{STV}$_{i}$, using proofs  established in \textbf{DeepSpec}, we instantiate the compiler to synthesise executable  verifier $\mathcal{V}_{i}$.  
\section{Generic STV Formalised}\label{sec:GenCertVer}
We proceed through three steps, namely specification, implementation and verification of implementations in HOL4 to formally verify the machine and its components.
\paragraph*{Specification}  
Recall that we require minimising the trusted computing base  for computing with generated verifiers. The specification module consists of properties which we can use to verify that our implementations of  STV algorithms in HOL4 indeed match with the description of the counting methods used.  
 In light of Definition~\ref{STVMachine} these properties   actually form the semantics of the generic STV machine.   
\paragraph*{Implementation} Our final objective for creating the
framework is to produce means for validating concrete items of evidence. To this end, we need decision procedures used for deciding whether or not  given evidence is valid. Therefore we define boolean-valued functions in the implementation sub-module to computationally realise logical declarations in the specification module. In particular, we implement decision procedures whose computational content  provably realises the specification of the machine semantics.  
\paragraph*{Verification} Here we formally prove that the implementations meet the expectations of their respective specifications. Therefore through verification we come to eliminate a trusted layer  required to lay in our framework.

@@

\item As the abstract syntactic representation of evidence closely represents concrete evidence, designing the parser and inspecting its correctness is less challenging.  
 \item HOL4 has well-developed libraries containing verified assertions of operations on lists. By using lists to represent the

@@

 tactics for discharging proof obligations on assertions involving list structure and operations on it. This significantly eases the process of verification. 

@@

There is also a reason for choosing the type $rat$  
%as the underlying arithmetic for the formalisation, 
instead of e.g. floating numbers. Elections with an STV counting scheme have a small margin of victory especially for the last vacancy left to fill \cite{MBlo}. This margin may happen to be less than the magnitude of the error caused by the accumulation of rounding errors as it is with floating points.  Hence one must sensibly avoid the high cost of electing a wrong candidate by falling into traps of imprecise calculations due to unintelligent choices. Using exact fractions allows safe handling of calculations. 
\subsection{The Semantics and its Auxiliary Components}
\label{sec:MachineSem}

The semantics of each machine transition label consists of conjunctions of formally specified general pre- and postconditions across STV schemes which enforce when and how to take a tallying action and what the immediate effect is. To demonstrate the process of specification, implementation and verification of the machine transitions and their semantics, we discuss the transfer-elected transition.

@@

%\footnote{The Tasmania STV also uses a similar surplus transfer mechanism.}.
 Instantiation of other machine transitions proceeds in a  similar manner. Under the section 'Step 3' and 'transfer surplus from elected candidates' the protocol explains under what conditions and how to distribute surplus votes. We summarise and rephrase these sections as follows. 

@@

\item[$\bullet_{6}$] only the last parcel of votes received (which resulted in a surplus) is transferred. It may be that the last parcel is the only parcel in a candidate's pile (if only one application of the count action has occurred), or more parcels exist (if several actions of count have happened as a result of an earlier  elect, transfer or eliminate actions)

@@

The lines denoted by the star symbol~($\star$) in TransferActSpec formalise the Clause~$\bullet_{7}$ above. Also note that we place the last parcel of the pile of the candidate $c$ whose votes are transferred first into the list of uncounted ballots so that in the subsequent transition count deals with distributing the votes and ``recalculating'' candidates' tallies and piles. Finally, because of efficiency purposes our system is designed to update the fractional transfer value of the surplus in the formalisation of the elect transition instead of the transfer-elected transition.

@@

For example, we represented the instantiation of transfer-elected with

@@

work. This is different in the actual proof engineering development.  We carry out different instantiations of the machine in separate modules but uniformly name instantiations of the transitions. For example, for instantiation of the machine with the ACT STV, we have a module accordingly named and inside the module we refer, for instance, to the instantiation of transfer-elected as TransferDec or TransferSpec without the infix Act.

@@

\emph{Scenario three.} 
 Suppose there is a clause in the description of $\mathcal{A}$ whose specification, and therefore implementation,  does not exist in the auxiliary module. Then one trivially has to extend the auxiliary module by specifying and implementing that clause and then verifying the implementation correct against the specification.  If one was overconfident in their implementation, then they can take the implementation as its specification in which case no verification is required.  Note that as including the generic formal machine transitions in the semantics of each transition instantiation is mandatory  
 there are few numbers of such clauses and consequently few lines of encoding needed to formalise them.

@@

 The function check\textunderscore{}count accepts a file as input on the command line, opens the file consisting of  evidence which we intend to validate, parses the header of the evidence consisting of the quota, number of seats and competing candidates in the election. If the header is well-formed, then check\textunderscore{}count proceeds to parse two judgement lines at a time and if the lines  successfully parse into  values of the type judgement (the type of the machine states) then checks if  the transition happening between them \emph{corresponds with the specification of the verifier in HOL4 ($\mathcal{V}_{spec}$)}. This process of reading evidence lines, parsing and checking them continues until either the evidence is accepted as valid or  check\textunderscore{}count encounters a malformed judgement line or an invalid transition step from one parsed judgement (machine state) to another  in which case it returns an error message  informing us where it occurs.

@@

Use of theorem proving to verify vote counting in elections
is an emerging field. The only use of theorem proving and certifying algorithms~\cite{CertAlg} that we are aware of is our own previous work~\cite{GhaleVSTTE} which just verifies a single instance of STV. While the overall approach is similar, the underlying data structure and the algorithmic content which is being formalised does not allow modular specification, verification and synthesis of verifiers for different types of STV: only a single STV algorithm is
being verified.  Adapting to other schemes
requires a considerable effort of re-encoding (sometimes the whole data structure) and non-trivial expertise with HOL4 and CakeML.

Other approaches focus on the  formalisation and verification of algorithms for correctly computing election results, rather than checking correctness of computation performed. This is usually based on so-called heavy-weight formal methods (interactive theorem proving) and uses a variety of 
%different
 tools.

There is work which formally verifies some STV algorithms in the theorem prover Coq. Ghale~et~al~\cite{Ghale:2017:FVS} formalise the student union STV algorithm used in some union elections Australia. Pattinson and Schurmann~\cite{DBLP:conf/ausai/PattinsonS15} verify the ``vanilla'' STV algorithm.   Pattinson and Tiwari~\cite{DBLP:conf/itp/PattinsonT17} prove the Schultz voting scheme correct, and Verity and Pattinson~\cite{DBLP:conf/acsw/VerityP17} verify an invariant of STV. Finally, Ghale~et~al~\cite{} provide a modular treatment of formalisation and verification of STV algorithms. They all then employ the automatic program extraction tool of Coq to obtain Haskell or OCaml programs to provably correctly compute election results based on the algorithm for any real input data. These programs output verifiable evidence upon each execution. Some of them~\cite{DBLP:conf/ausai/PattinsonS15,DBLP:conf/itp/PattinsonT17}    
 implement a verifier for checking the correctness of execution instances. However, the verifier is not verified. They implement their verifier as Haskell programs without proving that the verifier itself computes correctly. Moreover, even trusting that the verifier is reliable, there remain other unverified layers to trust in. For example, they rely on the GHC compiler to execute the program. As GHC is not a verified compiler one has to trust that the machine code executed also behaves correctly.
